---
title: "Lab07 ANOVA"
author: "Name"
date: "XX XX, 20XX"
output:
  html_document:
    toc: yes
  word_document:
    toc: yes
---

# Chunk 1 Load data and packages
  1.1 Load the 'tidyverse' package.

  1.2 Upload two csv files for this week's lab: 'HIVmutation.csv' and 'OpsinExpression.csv'.

  1.3 *describe the structure of each dataset* Which data set in appropriate for a 1-way ANOVA and which for a 2-way ANOVA?
  
```{r, warning = FALSE, message = FALSE}
library(tidyverse)

opsin = read.csv("OpsinExpression.csv", header = T)
str(opsin)

HIV = read.csv("HIVmutation.csv", header=T)

```

# Chunk 2 Make a QQ-plot (Normal probability plot) to visualize normality
Use the HIV dataset for the following instructions
  2.1 *Describe how to interpret a QQ plot / Normal probability plot?*

  2.2 Make two qq plots: one for the variable 'fitness' and another for 'logOddsFitness' which represent untransformed and transformed data, respectively.

  2.3 Add a ggtitle of "Untransformed" and "Transformed" to each of your QQ plots

  2.3 Compare the two qq plots. *State your conclusion.* Should you use the untransformed or transformed data for ANOVA?

```{r}
#How to interpret QQ plot:
?qqplot

#QQplots HIV data (with titles)
?ggtitle
ggplot (HIV, aes (sample=fitness)) + geom_qq() + ggtitle("Untransformed")
ggplot (HIV, aes (sample=logOddsFitness)) + geom_qq() + ggtitle("Transformed")

#Conclusion

```

# Chunk 3 ANOVA, Tukey-Kramer, and Krustkal Wallis tests
  3.1 Perform a Kruskal-Wallis test using untransformed 'fitness' as the dependent variable (HIV data). *Is there a significant effect of mutation type on fitness?*

  3.2 Write a null and alternative hypothesis for an ANOVA

  3.3 Perform an ANOVA using transformed 'logOddsFitness' as the dependent variable (HIV data). *Report the two df, F, and P values*. 

  3.4 State your conclusions about the ANOVA. *Is there a significant effect of mutation type on fitness?*

  3.5 If your ANOVA was significant, follow up with a Tukey-Kramer post-hoc test. *Which mutation types were significantly different from each other?*
```{r}
#Nonparametric test
kruskal.test(fitness ~ mutationType, data=HIV)

#Ho:
#Ha:  

#ANOVA 
aov1 = aov(logOddsFitness ~ mutationType, data=HIV)
summary (aov1)
#Conclusions:

#Tukey-Kramer Post-hoc comparisons
TukeyHSD (aov1)
plot (TukeyHSD (aov1))

#Post-hoc conclusions:

```

# Chunk 4  Bar and Error Plots
  4.1 Using pipes, summarize the HIV dataset so that you have a mean and standard error per group. 
  
  4.2 Add to the summary table the a back-transformed mean and upper and lower limits of the error bars. 

  4.3 Plot the back-transformed mean and error bars (HIV data) as a Bar and error plot.

  4.4 Rearrange the order of the categories of mutation type so that the order is 'synonymous', 'minorAAchange', 'majorAAchange', 
```{r}
#Back-transformed HIV mutation Bar and Error plot +/- se
hiv.summary<- HIV %>% group_by(mutationType) %>% 
        summarize (mean.logFitness = mean(logOddsFitness), 
              se.logFitness = sd(logOddsFitness/sqrt(n())),
                  upper = mean.logFitness + se.logFitness, 
                  lower = mean.logFitness - se.logFitness,
                  backmean = 2.718^mean.logFitness,
                  backup = 2.718^upper,
                  backlow = 2.718^lower)
hiv.summary

ggplot(hiv.summary, aes (x=mutationType, y=backmean)) + geom_col() + 
  geom_errorbar(aes(ymin=backlow, ymax=backup), width=0.1) +
  theme(panel.background = element_blank()) +
  theme(panel.border = element_rect(fill=NA)) +
  ylab("Ave fitness")

#rearranging order of categories
factmutation = factor(hiv.summary$mutationType, levels = c('synonymous', 'minorAAchange', 'majorAAchange'))
factmutation

ggplot(hiv.summary, aes (x=factmutation, y=backmean)) + geom_col() + 
  geom_errorbar(aes(ymin=backlow, ymax=backup), width=0.1) +
  theme(panel.background = element_blank()) +
  theme(panel.border = element_rect(fill=NA)) +    
  ylab("Ave fitness")
```

# Chunk 5: 2-way ANOVA
Use the Opsin data set for the following chunks:
  5.1 *Write the three null hypotheses that are being tested with a 2-way ANOVA*
  5.2 Perform a 2-way ANOVA.*Interpret the ANOVA source table*

# Chunk 6: Point and Error Plot of 2-way ANOVA.
  6.1 Using pipes, create separate 'Swamp' and 'Spring' summary tables for mean and se relative Expression of SWS1 gene (grouped by water clarity category.
  
  6.2. Combine the summary tables and add a column with the Population names.
  
  6.3  Make a point and error plot for the opsin data. Label the y axis "Relative expression of SWS1 gene" and x axis "Population". 
  
  6.4 Designate shape and color to "water quality levels" and increase the size of the points.
  
  6.5 *State your conclusions*. Which treatment groups is gene expression greater?

```{r}
#making the summary table
Spring = filter(opsin, population =="Spring") 
Swamp = filter(opsin, population =="Swamp")

#spring summary
spr.summary = Spring %>% group_by(waterClarity) %>% 
                  summarize (mean.exp = mean(relativeExpressionOfSWS1), 
                            se.exp = sd(relativeExpressionOfSWS1/sqrt(n())))

#Swamp Summary
swa.summary = Swamp %>% group_by(waterClarity) %>% 
                  summarize (mean.exp = mean(relativeExpressionOfSWS1), 
                            se.exp = sd(relativeExpressionOfSWS1/sqrt(n())))


combo.summary = rbind(spr.summary, swa.summary)
Population = c("Spring", "Spring", "Swamp", "Swamp"); 
Population = as.data.frame(Population)

total.summary = cbind(Population, combo.summary)  

#Point and error plot
ggplot(data = total.summary, aes(x=herbivores, y=mean.area, color = Height, shape = Height)) + geom_point(size=3) +
  geom_errorbar(aes(ymin = mean.area - se.area, ymax = mean.area + se.area), width = 0.1) +
  theme(legend.text=element_text(size=12)) +
  theme(panel.background = element_blank(), panel.border = element_rect(fill=NA)) +
  theme(axis.text=element_text(size=12)) + 
  ylab("Area of algal cover") +
  xlab("Hebivore Treatment")

#Fancy plot
ggplot(data = total.summary, aes(x=Population, y=mean.exp, color = waterClarity, shape = waterClarity)) + geom_point(size=3) +
  geom_errorbar(aes(ymin = mean.exp - se.exp, ymax = mean.exp + se.exp), width = 0.1)

#Conclusions
```
#knit and submit