---
title: "Lab 05- T-tests"
author: "YIsrael"
date: "CHANGE THIS "
output:
  word_document:
    toc: yes
  html_document:
    toc: yes
---
# Chunk 1: Load libraries and datasets 
  ## 1.1) Load `tidyverse` 
    ### THIRD LEVEL HEADER 
  ## 1.2) read in two datasets "WormSperm.csv" and "BlackbirdTesto.csv"
  ## 1.3) What are the structures of each of these two datasets? 
```{r, echo=FALSE}
setwd("/Users/israel/Desktop/Bio170_Spring22/Labs/Lab05- t-test")
library (tidyverse)
library (car)
sperm<- read_csv("WormSperm.csv")
bbirds<- read_csv("BlackbirdTesto.csv")
str (sperm)
```

# Chunk 2: Define Hypotheses and Plot the data! 
  2.1) Using pipes, summarize the bird and sperm datasets so that you have a mean and standard error per group (treatment in birds and sex in sperm). 
  2.2) Identify a null and alternative hypotheses testing for differences in means in each of these datasets 
  2.3) Plot summarized data as a point plot with error bars representing the standard error 
```{r, echo=FALSE}

bird.summary<- bbirds %>% group_by (Treatment) %>% 
  summarize (mean.antibody=mean(Antibody.production),
             se.antibody=sd(Antibody.production)/ sqrt (n()))

bird.summary

sperm.summary<- sperm %>% group_by(Sex) %>% 
  summarize (mean.size=mean (SpermSize), 
             se.size= sd(SpermSize)/sqrt (n()))

ggplot (bird.summary, aes (x=Treatment, y=mean.antibody)) + geom_point() + 
  geom_errorbar(aes(ymin=mean.antibody-se.antibody, ymax=mean.antibody+se.antibody), width=0.25)

ggplot (sperm.summary, aes (x=Sex, y=mean.size)) + geom_point() + 
  geom_errorbar(aes (ymin=mean.size-se.size, ymax=mean.size+se.size), width=.25)
```

# Chunk 3: Check statistical assumptions
  3.1) Using a shapiro test, a density plot and a qq plot check the two datasets for normality
    3.1b) If any dataset is not normally distributed attempt a transformation to normalize the data. Does the transformation improve the normality of the distribution? 
  3.2) Using both a Levene Test and a Flinger-Kileen test, check if the datasets meet the assumption of homeogeneity of variances 
  3.3) What about the assumption of independence, are all observations in each of the datasets independent of each other? 
  
```{r, echo=FALSE}
?shapiro.test
?car::leveneTest

#normality 
shapiro.test (bbirds$Antibody.production)

ggplot (bbirds, aes(Antibody.production, fill=Treatment)) + 
  geom_density(alpha=0.5) + xlim(0,200)


ggplot (bbirds, aes (sample=Antibody.production, color=Treatment)) + 
  geom_qq() + stat_qq_line()

bbirds$Antibody.log= log (bbirds$Antibody.production)
shapiro.test (bbirds$Antibody.log)

ggplot (bbirds, aes (Antibody.log, color=Treatment)) + geom_density()

shapiro.test (sperm$SpermSize)
ggplot (sperm, aes(SpermSize, fill=Sex)) + geom_density(alpha=0.5)

ggplot (sperm, aes (sample=SpermSize, color=Sex)) + geom_qq() + stat_qq_line()
sperm$log.size<- log (sperm$SpermSize)
shapiro.test(sperm$SpermSize)

#homogeneity of variance 
leveneTest(bbirds$Antibody.production, group=bbirds$Treatment)
leveneTest(sperm$SpermSize, group=sperm$Sex)

#alternative to levene test is Fligner-Kileen test for two groups
?fligner.test
fligner.test(Antibody.production ~ Treatment, bbirds)
fligner.test(SpermSize ~ Sex, sperm)

```

# Chunk 4: Compelte an unpaired t-test 
  4.1) Complete a t.test for both the transformed and untranfromed Antibody production predicted by Treatment in the bbirds dataset. Set the `var.equal=TRUE`
  
  4.2) Complete a t.test for both the transformed and untranfromed SpermSize predicted by Sex in the sperm dataset. Set the `var.equal=FALSE`
  
  4.3) Interpret the results in your own words. 
```{r, echo=FALSE}
?t.test
t.test (Antibody.production ~ Treatment, data=bbirds, var.equal=F)
t.test (Antibody.log ~ Treatment, data=bbirds, var.equal=TRUE)

t.test (SpermSize ~ Sex, data=sperm, var.equal=FALSE)
t.test (log.size ~ Sex, data=sperm, var.equal=FALSE)
```

# Chunk 5: Complete a paired t-test and plot a paired point plot 
  5.1) Complete a paired t test for the bbirds dataset. use `paired=TRUE`
  
  5.2) Interpret the results and discuss Why is this the appropriate test for these data? 
  
  5.3) Plot the data as paired point plot
  

```{r, fig.height=4, fig.width=6, res=600, echo=FALSE}
library (tidyverse)
?geom_point()
?geom_line()
str(bbirds)
t.test (Antibody.production ~ Treatment, data=bbirds, var.equal=TRUE, paired=T)
t.test (Antibody.log ~ Treatment, data=bbirds, var.equal=TRUE, paired=T)

ggplot (bbirds, aes(x=Treatment, y=Antibody.log)) + 
  geom_point(size=6, levels = c("Before", "After"), color="blue") +
  geom_line(aes(group=blackbird)) + theme_bw() + 
  theme (axis.text=element_text (size=18), 
         axis.title=element_text (size=18, face="bold") )
  
```

# Chunk 6: Non-parametric alternative 
6.1) Complete a wilcoxon rank sums test of Antibody.log predicted by treatment with equal variances and paired data. Interpret the results in your own words. 

6.2) Using the sperm data, complte a wilcoxon rank sums test of log.sperm size predicted by sex with unequal variances. Interpret the results in your own words. 
```{r, echo=FALSE}
wilcox.test(Antibody.log ~ Treatment, data=bbirds, var.equal=TRUE, paired=T)
ggplot (bbirds, aes(Treatment, y=Antibody.log)) + geom_boxplot()

wilcox.test (log.size ~ Sex, data=sperm, var.equal=FALSE)

```



