---
title: "Lab04: Chi Square and Power Analyses "
author: "YOUR NAME HERE!"
date: "XX/XX/20XX"
output: 
  html_document:
    toc: TRUE
---
#Chunk 1: Load libraries, data and set a working directory 
  1.1) Load `tidyverse` library 
  1.2) Load the "HumanHemoglobinElevation.csv" dataset and call it `Hemoglobin`
  1.3) What is the structure of this dataset? 
  1.4) What is the global mean of hemoglobin levels? 
  1.5) What is the population mean by country? Hint use `filter()`
  
```{r}
setwd("/Users/israel/Desktop/Bio170_Spring22/Labs/Lab04- Chi-sq")
#load libraries 
library (tidyverse)
#load data 
Hemoglobin<-read.csv("HumanHemoglobinElevation.csv", header = T)
# Structure 
?str
str (Hemoglobin)
# Summary 
?summary
summary (Hemoglobin)
#population mean
?filter
library (dplyr)
USA<-filter (Hemoglobin, population=="USA"); summary (USA); mean (USA$hemoglobin)
Andes<- filter (Hemoglobin, population=="Andes"); summary (Andes)
Ethiopia<- filter (Hemoglobin, population=="Ethiopia"); summary (Ethiopia)
Tibet<- filter (Hemoglobin, population=="Tibet"); summary (Tibet)
```

#Chunk 2: Plot the data
  2.1) Plot the distirbutions of hemoglobin levels by population as a density plot (set alpha to 0.5)
  2.2) Plot the distirbutions of hemoglobin levels by population as a boxplot
  2.3) Plot the distirbutions of hemoglobin levels by population as a strip plot inside a violin plot~!  (set alpha of points to 0.25 and color by population)
  2.4) Describe the different plots. What are the pros and cons of each way of visualizing distributions? 
    2.4 Answer: 
  
```{r}
ggplot (Hemoglobin, aes (x=hemoglobin, fill=population)) + geom_density(alpha=0.5)

ggplot (Hemoglobin, aes (x=population, y=hemoglobin, fill=population)) + geom_boxplot(alpha=0.5)

# THIS IS MY RESPONSE TO 2.4 
ggplot(Hemoglobin, aes(x=population, y=hemoglobin, color=population)) + 
  geom_violin(trim=F)+
  geom_jitter(position=position_jitter(0.2), alpha=0.25)

ggplot(Hemoglobin, aes(x=population, y=hemoglobin, color=population)) + 
  geom_violin(trim=T)+
  geom_jitter(position=position_jitter(0.1), alpha=0.25)
```

#Chunk 3: Plotting 2 categorical vairables 
  3.1) Create a stacked bar plot of the number of observations in each country by elevation category
  3.2)  Create a stacked bar plot of the percentage of observations in each country by elevation category
  
  3.3) How do these plots differ? 
  
```{r}
library (ggplot2)
ggplot (Hemoglobin, aes(population, fill=elevation)) + geom_bar()
ggplot (Hemoglobin, aes(population, fill=elevation)) + geom_bar(position="fill")
```

#Chunk 4: Chi-sq: 
  4.1) You are interested in the association between the number of participants in the study within each  country and elevation. What are the null and alternative hypotheses that can be tested? 
  
  4.2) create a contingency table of obervations in different populations vs. elevations
  
  4.3) create a proportional table of obervations in different populations vs. elevations
  
  4.4) calculate a chi-square test testing the association betweenthe number of participants in the study within each  country and elevation. 
  
  4.5) Interpret the results. Which of your hypotheses is supported? 
  
  4.6) repeat the procedure in 4.1-4.5 to test the association between elevation and hemoglobin levels? 
  
```{r}
?table
?prop.table
?chisq.test
  
table1<-table (Hemoglobin$population, Hemoglobin$elevation)
table1
prop.table(table1)

chisq.test(table1)

table2<-table (Hemoglobin$hemoglobin, Hemoglobin$elevation)
prop.table(table2)

chisq.test(table2)
```

#Chunk 5: Power Analysis 
  5.1) load the `pwr` library 
  5.2) Calculate the power for each of the four countries, given the proportions of observations and sample sizes for each country and a significance level of 0.05. Which countries have the highest power? Why do you think this is the case? 
  5.3) What happens when you decrease the expected significance level to 0.01. Interpret these results. 
  5.4) What happens when you decrease half the number of observations collected in each population but retain a significance level of 0.5. Interpret these results. 
```{r}
library (pwr)
?ES.h
?pwr.p.test
#Andes
h1<- ES.h(0.03618756, 0.00000000); h1
pwr.p.test(h=h1,n=171,sig.level=0.05,alternative="two.sided")

#Ethiopia

h2<- ES.h(0.04892966,0.01630989); h2
pwr.p.test(h=h2,n=128,sig.level=0.05,alternative="two.sided")


#Tibet 
h3 <- ES.h(0.03007136, 0.00000000); h3
pwr.p.test(h=h3,n=59,sig.level=0.05,alternative="two.sided")

#USA 
h4 <- ES.h(0.00000000, 0.86850153); h4

pwr.p.test(h=h4,n=1704,sig.level=0.05,alternative="two.sided")
```

# Chunk 6: Fancy Plotting

  6.1) Using pipes create a summary showing the means, number of obervations per country, standard deviations and standard errors for the levels of Hemoglobin in each country. 
  6.2) Calculate a 95% Confidence interval for each country's hemoglobin level 
  6.3) Plot a point plot (representing the mean) with error bars tha that show the 95%CI 
  6.4) Plot a point plot (representing the mean) with error bars tha that show the standard error
  6.5) Plot a point plot (representing the mean) with error bars tha that show the standard deviation
  
```{r}
?group_by
?summarise
?qnorm

summary1<-Hemoglobin %>% group_by(population) %>% 
  summarize (mean.hemo = mean (hemoglobin), 
             n.hemo = n(),
             sd.hemo = sd(hemoglobin),
             se.hemo= sd(hemoglobin)/sqrt(n())) 

summary1

zstar<- qnorm(0.025, lower.tail = FALSE); zstar
zstar<- qnorm(0.005, lower.tail = FALSE); zstar
zstar<- qnorm(0.05, lower.tail = FALSE); zstar

CI95<- summary1 %>% mutate (upper.95= mean.hemo + se.hemo*zstar, 
                            lower.95= mean.hemo - se.hemo*zstar)

CI95

ggplot (CI95, aes(x=population, y=mean.hemo, color=population)) + geom_point() + 
   ylim(0,20) + geom_errorbar(aes(ymin=lower.95, ymax=upper.95), width=0.1)

```