---
title: "Lab 3- Distributions"
author: "COMPLETE KEY"
date: "XX/XX/20XX"
output:
  html_document:
    toc: yes
---
#Chunk 1: Load libraries and Data
	1) Load the packages “ggplot2”, “tidyverse" and “datasets” 
	2) Load the dataset “sleep” 
	3) Using a comment describe the structure `str()` of the ‘sleep’ dataset

```{r, echo=FALSE}
library (tidyverse)
library (datasets)
data(sleep)
str (sleep)
```

#Chunk 2: One Dimmensional Data Visualization- Histograms
  1) Using ‘ggplot’, ‘geom_histogram()’, ‘gg_title’ plot a histogram of the sleep dataset’s variable “extra” using the default bins setting and title each plot based on the number of bins displayed.

  2) Now modify the `bins` argument from 30 (default) and create new plots, (change the title to reflect the number of bins being displayed) 
	bins=5
	bins=10
	bins=20
	bins=50

  3) Using a comment describe what happens when you change the bin size? 

```{r, echo=FALSE}
?ggplot()
?geom_histogram()
?ggtitle()

ggplot (sleep, aes (extra)) + geom_histogram(bins=50) + ggtitle ("bins=50")
ggplot (sleep, aes (extra)) + geom_histogram(bins=30) + ggtitle ("bins=30")
ggplot (sleep, aes (extra)) + geom_histogram(bins=10) + ggtitle ("bins=10")
ggplot (sleep, aes (extra)) + geom_histogram(bins=5) + ggtitle ("bins=5")
```

#Chunk 3: One Dimmensional Data Visualization- Density Plots
  	1) Create a density plot for the variable “extra” in the “sleep” dataset	

	  2) modify the x axis to range from -10 to 10

	  3) Create a new density plot but color the density plot by “group” with a 50%   transparency 

```{r, echo=FALSE}
?ggplot()
?geom_densiyty
ggplot (sleep, aes (extra)) + geom_density() + xlim(c(-10,10))

ggplot (sleep, aes (extra, fill=group)) + geom_density(alpha=.5) + xlim(c(-10,10))
```

#Chunk 4: Generate a Random Normal Distribution
	1) Create an object called “extra” of a normal distribution with a 20 observations a mean of 0 and a standard deviation of 1. 

	2) Convert norm from a list to a data.frame (?as.data.frame)

  3) Plot it as a density plot 


```{r, echo=FALSE}
set.seed(198)
?rnorm 
?as.data.frame

extra <- rnorm(120, 0, 1)
extra<-as.data.frame (extra)

ggplot (extra, aes (extra)) + geom_density() + xlim (c(-10,10))

shapiro.test (extra$extra)

```

#Chunk 5: Data Merging and plotting

5.4 Change the number of observations to 10 and plot it as a density plot. Using a 
comment, how does the distribution change? 

5.5. Change the number of observations to 100 and plot it as a density plot. Using a 
comment, how does the distribution change? 

5.6 Use `shapiro.test` to test for normality in this distribution? What do the results suggest? Are the data normally distributed? H0 represents that the data are not different from a normal distribution. 
```{r, echo=FALSE}
?add_column
?rbind

norm.table<- add_column (extra, group="RANDOM", ID="RANDOM")

new.df<-rbind (norm.table, sleep)
str (new.df)

ggplot (new.df, aes (extra, fill=group)) + geom_density(alpha=.5) + xlim(c(-10,10))

group1<-filter (sleep, group=="1")
shapiro.test(group1$extra)

group2<-filter (sleep, group=="2")
shapiro.test(group2$extra)
```
# Chunk 6: Check for Kurtosis and Skewness 
```{r}
library (e1071)
?kurtosis
?skewness
#Normal Distribution
kurtosis(extra$extra, type = 2)
skewness( extra$extra, type = 2)

kurtosis(group1$extra, type = 2)
skewness( group1$extra, type = 2)

kurtosis(group2$extra, type = 2)
skewness( group2$extra, type = 2)

```

#Chunk 7: Fancy Plotting
Do any five of these things to the version of the plot completed in chunk 5:
	
	* change the default color of the plot to different color using  ‘scale_fill_manual’
	
	* change the x label to “Effect on hours of sleep”
	
	* change the background theme to black and white ‘theme_bw()’
	
	* add a vertical line at the mean value ‘geom_vline()’
	
	* change the label on the color key ‘labs’
	
	* change the position of the key from the right to the top or bottom 
	
	* add a title to the plot ‘ggtitle’
	
	* annotate the plot with some text (e.g. describing the vertical line) ‘annotate’
	
	* change the figure size to 4 in X 6 in (markdown chunk modification)
	
	* change the output resolution of the figure to 400 dpi (markdown chunk modification)
	
```{r, fig.height=4, fig.width=6, res=600, echo=FALSE}
summary (new.df)

ggplot (new.df, aes (extra, fill=group)) + geom_density(alpha=.5) + 
  xlim(c(-10,10)) + scale_fill_manual(values=c("Grey", "DarkGrey", "Black")) +   
  theme_bw() + xlab("Effect on Hours of Sleep") + 
  labs(fill = "Treatment Group") + geom_vline(xintercept=0.9004) + 
  annotate ("text", x=7, y=.35, label= "distribution mean=0.9", size=10) +
  theme (legend.position = "top", legend.text = element_text(size=18), 
         legend.title = element_text(size=18, face="bold"), 
         axis.text=element_text(size=18), axis.title=element_text(size=18, face="bold"))
```


